A Methodological Approach Comparison to Estimate Mortality in Small Areas: the Pampean Region in Argentina (2009-2011)

### 2020-02-05 analysis/Paper_draft_ENG.Rmd
mas de la intro

---
title: "A Methodological Approach Comparison of Mortality Estimates for Small Areas: the Pampean Region in Argentina (2009-2011)"
author: |
  Nicolás Sacco [^ns]; Iván Williams [^iw]; Bernardo L. Queiroz [^bq]

@@

Do mortality disparities subsist among small-areas? Previous research in the demography of mortality differentials argue that they would be gradually dismiss by the epidemiological transition, public health policies, and improvements in quality of life (CITA). In most countries, life expectancy at birth ($q_0$) increase during the early 20th century, due the advancement of knowledge and medical technology to fight infectious diseases and improvements in living conditions associated with socioeconomic development.

Althought socioeconomic mortality differentials contributed to sustain the hypotesis of persistent inequality at death during and after the epidemiological transition (CITA, Palloni, ?), some authors expect that health outcomes and disparities would be extinguish as the demographic transition unfolds (CITA). Retrospective views, however, show that mortality differentials are persitent and have grown over time in many cases by socioeconomic status (CITA). Anyhow, it is useful account if this disparities continue and/or arise regardless of auspicious initial scenarios.

@@

Argentina represents a clear example of a country with few data sources, but also a very interesting case of the epidemiological transition, often not address by the previous literature. In this sense the goal of this article

----
 Often data in small areas are non-existent or sparce in this region. Spatial patterns are essential to understand individual demographic outcomes related with the characteristics of a place, also as a tool for the application of development plans and for the allocation of resources.

>>>>

---
title: "A Methodological Approach Comparison to Estimate Mortality in Small Areas: the Pampean Region in Argentina (2009-2011)"
author: |
  Nicolás Sacco [^ns]; Iván Williams [^iw]; Bernardo L. Queiroz [^bq]

@@

Do mortality disparities subsist among small-areas? Previous research in the demography of mortality differentials argue that they would be gradually dismiss by the epidemiological transition, public health policies, and improvements in quality of life (CITA). In most countries hig-income and middle-income countries, life expectancy at birth ($q_0$) increase during the early 20th century, due the advancement of knowledge and medical technology to fight infectious diseases and improvements in living conditions associated with socioeconomic development.

Althought socioeconomic mortality differentials contributed to sustain the hypotesis of persistent inequality at death during and after the epidemiological transition (CITA, Palloni, ?), some authors expect that health outcomes and disparities would be extinguish as the demographic change unfolds (CITA). Retrospective views, however, show that mortality differentials are persitent and have grown over time in many cases by socioeconomic status (CITA). Anyhow, it is useful account if this disparities continue and/or arise regardless of auspicious initial scenarios.

@@

On this subject, many recent contributions develop methods to estimate small area data. *****Mas sobre aportes recientes para estimacion de mortalidad en areas menores en el mundo.***.

Argentina represents a clear example of a country with few data sources, but also a very interesting case of the mortality transition, often not address by previous research. With all this, the aim of this article is to generate mortality estimates that exemplify the relationship between status and health. For that, we apply three different methods to estimate and then compare mortality levels in small areas in the Argentinean Pampean Region, during the period 2009-2011. To test how the choosen methods perform, we discriminate this region because is most populate region in the country.

We use administrative death register data from the Department of Health Statistics and Information of the Ministry of Health of the Nation (DEIS, acronym in Spanish; the government offical unit in charge of death counts administration), and population estimates prepare by the National Institute of Statistics and Census (INDEC, acronym in Spanish; the Argentine government agency responsible for the collection and processing of statistical data, such as population censuses). Using three different approaches, we first apply Death Distribution Methods (DDM) to analyze We then combine DDM with empirical bayesian approach to produce mortality estimates for small areas (municipalities) in Brazil. The two-step procedure, allows us to correct the underegistration of death counts and minimize random fluctuations that might occur when we calculate mortality levels in small areas.

Debido a la poca experiencia de estudios previos que estimen mortalidad *general* en áreas menores en Argentina, se decidió aplicar tres técnicas: una basada en la teoría bayesiana, la segunda basada en métodos relacionales de tablas de vida pero agregando técnicas estadísticas de suavizado, y tercero, un enfoque demográfico clásico, considerado el método por default debido a su simplicidad. Antes de la estimación, se realizó un procedimiento de regionalización para aprovechar la similitud espacial entre áreas pequeñas, independientemente de su pertenencia político-administrativa.

the quality of mortality data and estimate adult mortality levels for microregions in Brazil.

# Why Argetina? Why the Pampas?

Debido a la poca experiencia de estudios previos que estimen mortalidad *general* en áreas menores en Argentina, se decidió aplicar tres técnicas
